### 2023-09-06 proofs/tex/upgrade-lemmas.tex
Add deterinicity lemmas, get rid of (??)

\begin{lemma}[Soundness of Upgrade]\label{lemma:upgrade-soundness}
    Assuming $[[iP]]$ is well-formed in $[[Γ = Δ, pnas]],$\\
    if $[[upgrade Γ ⊢ iP to Δ = iQ]]$
    then
    \begin{enumerate}
        \item $[[Δ ⊢ iQ]]$
        \item $[[Γ ⊢ iQ ≥ iP]]$
    \end{enumerate}
\end{lemma}
\begin{proof}
    By inversion, $[[upgrade Γ ⊢ iP to Δ = iQ]]$ means that 
    for fresh $[[pnbs]]$ and $[[pncs]]$,
    $[[Δ, pnbs, pncs ⊨ [pnbs/pnas]iP ∨ [pncs/pnas]iP = iQ]]$.
    Then by the soundness of the least upper bound (\cref{lemma:lub-soundness}),
    \begin{enumerate}
        \item $[[Δ, pnbs, pncs ⊢ iQ]]$, 
        \item $[[Δ, pnbs, pncs ⊢ iQ ≥ [pnbs/pnas]iP]]$, and 
        \item $[[Δ, pnbs, pncs ⊢ iQ ≥ [pncs/pnas]iP]]$.
    \end{enumerate}

    $ 
    \begin{aligned}
        [[fv iQ]] &\subseteq [[fv [pnbs/pnas]iP ∩ fv [pncs/pnas]iP]]
                  &&\text{Since by \cref{lemma:fv-propagation}, 
                         $[[fv iQ ⊆ fv [pnbs/pnas]iP]]$ and
                         $[[fv iQ ⊆ fv [pncs/pnas]iP]]$}\\
                  &\subseteq [[ ((fv iP \ {pnas}) ∪ {pnbs}) ∩ ((fv iP \ {pnas}) ∪ {pncs})]]\\
                  &= [[ (fv iP \ {pnas}) ∩ (fv iP \ {pnas}) ]]
                  &&\text{since $[[pnbs]]$ and $[[pncs]]$ are fresh}\\
                  &= [[ fv iP \ {pnas} ]]\\
                  &\subseteq [[ Γ \ {pnas} ]]
                  &&\text{since $[[iP]]$ is well-formed in $[[Γ]]$}\\
                  &\subseteq [[ {Δ} ]]\\
    \end{aligned}
    $\\
    This way, by \cref{lemma:wf-ctxt-equiv}, $[[Δ ⊢ iQ]]$.

@@

    This way, $[[iQ']]$ is a common supertype of $[[ [pnbs/pnas]iP ]]$ and $[[ [pncs/pnas]iP ]]$ in
    context $[[Δ, pnbs, pncs]]$. It means that we can apply the completeness of the least upper bound
    (\cref{lemma:lub-completeness}):
    \begin{enumerate}
        \item there exists $[[iQ]]$ s.t. $[[Γ ⊨ [pnbs/pnas]iP ∨ [pncs/pnas]iP = iQ]]$ 
        \item $[[Γ ⊢ iQ' ≥ iQ]]$.
    \end{enumerate}
    The former means that the upgrade algorithm terminates and returns $[[iQ]]$.
    The latter means that since both $[[iQ']]$ and $[[iQ]]$ are well-formed in $[[Δ]]$,
    by \cref{tood}, $[[Δ ⊢ iQ' ≥ iQ]]$.
\end{proof}

>>>>

\begin{observation}[Upgrade is deterministic]
    \label{obs:upgrade-deterministic}
    Assuming $[[iP]]$ is well-formed in $[[Γ ⊆ Δ]],$\\
    if $[[upgrade Γ ⊢ iP to Δ = iQ]]$ and $[[upgrade Γ ⊢ iP to Δ = iQ']]$ are defined 
    then $[[iQ = iQ']]$.
\end{observation}
\begin{proof}
    It follows directly from \cref{obs:lub-deterministic},
    and the convention that the fresh variables are chosen by a fixed deterministic algorithm
    (\cref{sec:fresh-selection}).
\end{proof}

\begin{lemma}[Soundness of Upgrade]\label{lemma:upgrade-soundness}
    Assuming $[[iP]]$ is well-formed in $[[Γ = Δ, pnas]]$,
    if $[[upgrade Γ ⊢ iP to Δ = iQ]]$
    then
    \begin{enumerate}
        \item $[[Δ ⊢ iQ]]$
        \item $[[Γ ⊢ iQ ≥ iP]]$
    \end{enumerate}
\end{lemma}
\begin{proof}
    By inversion, $[[upgrade Γ ⊢ iP to Δ = iQ]]$ means that 
    for fresh $[[pnbs]]$ and $[[pncs]]$,
    $[[Δ, pnbs, pncs ⊨ [pnbs/pnas]iP ∨ [pncs/pnas]iP = iQ]]$.
    Then by the soundness of the least upper bound (\cref{lemma:lub-soundness}),
    \begin{enumerate}
        \item $[[Δ, pnbs, pncs ⊢ iQ]]$, 
        \item $[[Δ, pnbs, pncs ⊢ iQ ≥ [pnbs/pnas]iP]]$, and 
        \item $[[Δ, pnbs, pncs ⊢ iQ ≥ [pncs/pnas]iP]]$.
    \end{enumerate}

    $ 
    \begin{aligned}
        [[fv iQ]] &\subseteq [[fv [pnbs/pnas]iP ∩ fv [pncs/pnas]iP]]
                  &&\text{since by \cref{lemma:fv-propagation}, 
                         $[[fv iQ ⊆ fv [pnbs/pnas]iP]]$ and
                         $[[fv iQ ⊆ fv [pncs/pnas]iP]]$}\\
                  &\subseteq [[ ((fv iP \ {pnas}) ∪ {pnbs}) ∩ ((fv iP \ {pnas}) ∪ {pncs})]]\\
                  &= [[ (fv iP \ {pnas}) ∩ (fv iP \ {pnas}) ]]
                  &&\text{since $[[pnbs]]$ and $[[pncs]]$ are fresh}\\
                  &= [[ fv iP \ {pnas} ]]\\
                  &\subseteq [[ Γ \ {pnas} ]]
                  &&\text{since $[[iP]]$ is well-formed in $[[Γ]]$}\\
                  &\subseteq [[ {Δ} ]]\\
    \end{aligned}
    $\\
    This way, by \cref{lemma:wf-ctxt-equiv}, $[[Δ ⊢ iQ]]$.

@@

    This way, $[[iQ']]$ is a common supertype of $[[ [pnbs/pnas]iP ]]$ and $[[ [pncs/pnas]iP ]]$ in
    context $[[Δ, pnbs, pncs]]$. It means that we can apply the completeness of the least upper bound
    (\cref{lemma:lub-completeness}):
    \begin{enumerate}
        \item there exists $[[iQ]]$ s.t. $[[Γ ⊨ [pnbs/pnas]iP ∨ [pncs/pnas]iP = iQ]]$ 
        \item $[[Γ ⊢ iQ' ≥ iQ]]$.
    \end{enumerate}
    The former means that the upgrade algorithm terminates and returns $[[iQ]]$.
    The latter means that since both 
    $[[iQ']]$ and $[[iQ]]$ are well-formed in $[[Δ]]$ and $[[Γ]]$,
    by \cref{lemma:subt-ctxt-irrelevance}, $[[Δ ⊢ iQ' ≥ iQ]]$.
\end{proof}
